---
title: "Problem Set 2"
author: "Kongpob Lee"
date: "2023-01-21"
output: pdf_document
---

################################################################################
##
## [ PROJ ] < Problem set 2 >
## [ FILE ] < Problem set 2 >
## [ AUTH ] < Kongpob Leemingsawat | Kongpob13 >
## [ AUTH ] < Shun Imazu | shunimazu>
## [ AUTH ] < Michael Rozen | michaelrozen3>
## [ INIT ] < 1/21/2023 >
##
################################################################################

## ---------------------------
## libraries
## ---------------------------

```{r}
library(tidyverse)
library(readxl)
library(labelled)
```

## ---------------------------
## directory paths
## ---------------------------

```{r}
setwd("/Users/gongpobleemingsawat/Desktop/MPP/First_Year/Winter_2023/EDUC_260B/PS2_Leemingsawat_Kongpob")

data_dir <- file.path('.', 'data')
plot_dir <- file.path('.', 'output', 'plots')
file_dir <- file.path('.', 'output', 'files')
```

## -----------------------------------------------------------------------------
## Part I: Command line & setting up git repo
## -----------------------------------------------------------------------------

# Question 3: 

- Command to change directories

pwd 
cd desktop
ls 
cd my_folder

- Command to clone repo

git clone https://github.com/anyone-can-cook/PS2_Leemingsawat_Kongpob.git

# Question 5: 

- Command to list directory contents, including hidden files/directories

ls -a 

- Output 

.               ..              .git            .gitignore      Problemset2.Rmd README.md

# Question 6: 

- Command to create new directories

mkdir -p data output/files scripts 
cd output 
mkdir plots 

- Command to create R script

cd scripts 
touch ps2_script.R

*to set path: cd desktop/MPP/First_Year/Winter_2023/EDUC_260B/PS2_Leemingsawat_Kongpob

## -----------------------------------------------------------------------------
## Part II - Working with data in R
## -----------------------------------------------------------------------------

# Question 2: 

```{r}
# Download data
download.file(url = 'https://nces.ed.gov/ipeds/datacenter/data/HD2019.zip',
              destfile = file.path(data_dir, "ps2_file.zip"))

# Unzip file 

unzip(zipfile =  file.path(data_dir, "ps2_file.zip"),
      exdir = data_dir)
```

# Question 3: 
```{r}
hd <- read_csv(file.path(data_dir, "hd2019.csv"),
               col_types = cols_only(UNITID = col_character(),
                                     INSTNM = col_character(),
                                     STABBR = col_character(),
                                     LONGITUD = col_double(),
                                     LATITUDE = col_double()
                                     ))
```

# Question 4: 

```{r}
# Filter 

hd_ca <- hd %>% filter(hd$STABBR == "CA")

# Plot 

png(file.path(plot_dir, 'ca_univs.png'))
ggplot(hd_ca, aes(x = LATITUDE, y = LONGITUD)) +
  geom_point() +
  theme_minimal() +
  coord_fixed(ratio = 1.5)
dev.off()
```

# Question 5: 

```{r}
write_csv(hd_ca, file = file.path(file_dir, "hd2019_ca.csv"))
```

## -----------------------------------------------------------------------------
## Part III - Git & GitHub 
## -----------------------------------------------------------------------------

# Question 1: 

Answer: git status 

Output: 'Untracked files'
.
.
# Question 2: 

answer: git add 

output: change to be committed 
.
.
# Question 3: 

answer: git commit -m data output scripts 

# Question 4: 

.
.
.

# Question 5: 

1. What heading is ‘.gitignore‘ listed under?

Changes not staged for commit

2. Command to show changes made to ‘.gitignore‘

git diff -- .gitignore

3. Output 

# diff --git a/.gitignore b/.gitignore
index fae8299..6154f65 100644
--- a/.gitignore
+++ b/.gitignore
@@ -1,3 +1,7 @@
+#Problem set files
+problemset2.md
+/.DS_Store
.
.
.
# Question 6: 

git add .gitignore PS2_Leemingsawat_Kongpob
.
.
.
# Question 7: 

1. command

git log 
.
.
.
# Question 8: 

1. command

git remote -v 

2. output

origin https://github.com/anyone-can-cook/PS2_Leemingsawat_Kongpob/ (fetch)
origin https://github.com/anyone-can-cook/PS2_Leemingsawat_Kongpob/ (push)
.
.
.
# Question 9: 

git push 

## -----------------------------------------------------------------------------
## Part III - GitHub Issue 
## -----------------------------------------------------------------------------

1. Paste the url to your issue here:
https://github.com/anyone-can-cook/rclass2_student_issues_w23/issues/96#issue-1558853048
2. Paste the url to the issue you responded to here:
https://github.com/anyone-can-cook/rclass2_student_issues_w23/issues/84#issuecomment-1405736210

## -----------------------------------------------------------------------------
## END SCRIPT
## -----------------------------------------------------------------------------